---
title: "Lab2_Co2_1997"
author: "Qiong Zhang"
date: "2024-03-10"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE, echo=F}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork) 

library(lubridate)
library(datasets)

library(tsibble)
library(feasts)
library(forecast)

library(sandwich)
library(lmtest)
library(ggfortify)
library(nycflights13)
library(blsR)
library(ggplot2)
```

```{r set themes}
theme_set(theme_minimal())
```

# Report from the Point of View of 1997 

For the first part of this task, suspend reality for a short period of time and conduct your analysis from the point of view of a data scientist doing their work in the early months of 1998. Do this by using data that is included in _every_ R implementation, the `co2` dataset. This dataset is lazily loaded with every R instance, and is stored in an object called `co2`. 

## (3 points) Task 0a: Introduction 

Introduce the question to your audience. Suppose that they _could_ be interested in the question, but they don't have a deep background in the area. What is the question that you are addressing, why is it worth addressing, and what are you going to find at the completion of your analysis. Here are a few resource that you might use to start this motivation. 

- [Wikipedia](https://en.wikipedia.org/wiki/Keeling_Curve)
- [First Publication](./background/keeling_tellus_1960.pdf)
- [Autobiography of Keeling](./background/keeling_annual_review.pdf)

## (3 points) Task 1a: CO2 data
Conduct a comprehensive Exploratory Data Analysis on the `co2` series. This should include (without being limited to) a [description of how, where and why ](https://gml.noaa.gov/ccgg/about/co2_measurements.html) the data is generated, a thorough investigation of the trend, seasonal and irregular elements. Trends both in levels and growth rates should be discussed (consider expressing longer-run growth rates as annualized averages).

```{r}
# Load the CO2 dataset
data("co2")

# View the structure of the CO2 dataset
str(co2)

# Create a dataframe with co2 and date columns
co2_df <- data.frame(co2)
dates <- seq(as.Date("1959-01-01"), by = "month", length.out = length(co2))

# Format the dates to extract the month and year in "YYYY-MM" format
co2_df$monthyear <- format(dates, "%Y-%m")

# Convert to ts series
co2_df <- co2_df %>% 
  mutate(time_index = yearmonth(monthyear)) %>% # convert the date to year-month format
  as_tsibble(index = time_index) # create time series 

head(co2_df)
```
```{r, fig.width = 7, fig.height = 5, fig.align = 'center', warning = F}
# Time plot
timeplot_co2 <- co2_df %>%
  ggplot() +
  aes(x=time_index, y=co2) +
  geom_line() +
  labs(
    title = "Monthly Mean CO_2",
    x = "Month and Year",
    y = "CO_2 Parts per Million"
  )

# Histogram
hist_co2 <- co2_df %>%
  ggplot() +
  geom_histogram(aes(x = co2)) +
  labs(
    title = "Histogram of CO_2",
    x = "CO_2 Parts per Million"
  ) +
  theme(legend.position = c(.2, .8))

# ACF Plot
acf_co2 <- co2_df%>%
  ACF(y=co2) %>%
  autoplot()

# PACF Plot
pacf_co2 <- co2_df %>%
  ACF(y=co2, type = "partial") %>%
  autoplot() +
  labs(y = "PACF")

(timeplot_co2 + hist_co2) /
  (acf_co2 + pacf_co2)
```
**Interpretation**:\
We observed strong and persistent trend and seasonality in monthly mean CO_2. The ACF plots fails to converge to 0. The histogram is not strongly skewed and we can observe the linear trend in the time plot, so we think log transformation is not necessary in this case.
```{r, fig.width = 7, fig.height = 5, fig.align = 'center', warning = F}
# First differencing of co2
timeplot_co2_1d <- co2_df %>%
  ggplot() +
  aes(x=time_index, y=difference(co2)) +
  geom_line() +
  labs(
    title = "1st Diff Monthly Mean CO_2",
    x = "Time Index",
    y = "1st Diff of Monthly Mean CO_2"
  )

# First differencing histogram
hist_co2_1d <- co2_df %>%
  ggplot() +
  geom_histogram(aes(x = difference(co2))) +
  labs(
    title = "1st Diff CO_2 Histogram",
    x = "1st Diff CO_2"
  ) +
  theme(legend.position = c(.2, .8))

# First differencing ACF
acf_co2_1d <- co2_df %>%
  ACF(y=difference(co2)) %>%
  autoplot()

# First differencing PACF
pacf_co2_1d <- co2_df %>%
  ACF(y=difference(co2), type = "partial") %>%
  autoplot()

(timeplot_co2_1d + hist_co2_1d) /
  (acf_co2_1d + pacf_co2_1d)
```
**Interpretation**:\
The time plot of the first differencing of monthly CO_2 indicates strong seasonality, which is further verified by the periodic oscillations in ACF plot. 
```{r, fig.width = 7, fig.height = 5, fig.align = 'center', warning = F}

dcmp_co2 <- decompose(co2)
dcmp_co2_1d <- decompose(diff(co2))

plot(dcmp_co2)
plot(dcmp_co2_1d)
```
**Interpretation**:\
Since the magnitude of the seasonal fluctuations, or the variation around the trend-cycle, does not vary with the level of the time series, we applied the additive decomposition.\
Based on the decomposition plot, we found:\
1) The trend of monthly mean CO_2 is linear.\
2) The seasonal fluctuations of monthly mean CO_2 does not change with the level of the time series.\
3) The level of the first-differencing monthly mean CO_2 does not vary much over time.
```{r, fig.width = 7, fig.height = 5, fig.align = 'center', warning = F}
co2_df$trend <- dcmp_co2$trend
co2_df %>%
autoplot(co2, color = "grey") +
geom_line(data = co2_df, 
          aes(x = time_index, y = trend, color = "Trend"))+
  labs(title = expression(Trend~of~Monthly~Mean~CO[2]), 
       y = "CO2 Parts per Million",
       x = "Month and Year")
```


What you report in the deliverable should not be your own process of discovery, but rather a guided discussion that you have constructed so that your audience can come to an understanding as succinctly and successfully as possible. This means that figures should be thoughtfully constructed and what you learn from them should be discussed in text; to the extent that there is _any_ raw output from your analysis, you should intend for people to read and interpret it, and you should write your own interpretation as well. 

## (3 points) Task 2a: Linear time trend model

Fit a linear time trend model to the `co2` series, and examine the characteristics of the residuals. Compare this to a quadratic time trend model. Discuss whether a logarithmic transformation of the data would be appropriate. Fit a polynomial time trend model that incorporates seasonal dummy variables, and use this model to generate forecasts to the year 2020. 

## (3 points) Task 3a: ARIMA times series model 

Following all appropriate steps, choose an ARIMA model to fit to the series. Discuss the characteristics of your model and how you selected between alternative ARIMA specifications. Use your model (or models) to generate forecasts to the year 2022. 

## (3 points) Task 4a: Forecast atmospheric CO2 growth 

Generate predictions for when atmospheric CO2 is expected to be at [420 ppm](https://research.noaa.gov/article/ArtMID/587/ArticleID/2764/Coronavirus-response-barely-slows-rising-carbon-dioxide) and 500 ppm levels for the first and final times (consider prediction intervals as well as point estimates in your answer). Generate a prediction for atmospheric CO2 levels in the year 2100. How confident are you that these will be accurate predictions?

